---
title: "Data Vizualizations Portfolio"
author: "Karlena Ochoa"
date: "February 24, 2019"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r load libraries }
 
library(tidyverse)
library(janitor)
library(rio)
library(magrittr)
library(here)
library(knitr)
library(stats)
library(gghighlight)
#devtools::install_github("yixuan/showtext")
library(showtext)
library(ggthemes)
library(grid)

```


```{r import data}

df <- import(here("data", "BREIF_Data_100p_final_proj.csv"),
             na.string = "999")
             
df2 <- import(here("data", "data_trust_final_proj.csv"),
              na.string = "999")

# turn 999 into nas

#join data and select variables of interest

data <- left_join(df2, df, by = "subid") %>% 
  select(1, 4:7, 12:23, 25:28, 33:44, 56:73, 81:143)
          
```


Viz 1: Aimed at general community!

Message: only 4-year-olds trust the reliable informant over the unreliable informant, even when 3-year-olds receive first-hand information they are equally choosing the reliable (helpful) and unreliable informant (tricky) in a sticker-finding task. 
```{r viz 1}

#mutate variables to factors and label, make average trust variable
v1 <- data %>% 
mutate(condition = as.factor(ifelse(condition == 1, "first-hand",
                                      "second-hand"))) %>%
  mutate(gender = as.factor(ifelse(gender == 1, "male", "female"))) %>%
  group_by(subid) %>%
  mutate(average_trust = mean(total_trust))

summary_p <- v1 %>%
  group_by(agegroup.x, condition) %>%
  summarise(n = n(),
            average_trust = mean(total_trust),
            sd_average_trust = sd(total_trust),
            se_average_trust = sd_average_trust / (sqrt(n)))

plot0 <- ggplot(summary_p, aes(as.factor(agegroup.x), average_trust )) +
  geom_bar(aes(fill = condition), stat = "identity", position = "dodge",
           alpha = .7) +
    labs(title= "Children's Selective Trust of Helpful over Tricky informant", 
       subtitle = "Separated by Condition and Age group", 
       x = "Age Group", y = "Average Trust Score" ) +
     geom_errorbar(aes(ymin = average_trust - 1.96*se_average_trust, 
                       ymax = average_trust + 1.96*se_average_trust,
                       group = condition),
                  width=.2,                    
                  position = position_dodge(width = 0.9)) +
  theme_minimal()

#improve plot by changing colors, make text bigger 

plot1 <- ggplot(summary_p, aes(as.factor(agegroup.x), average_trust )) +
  geom_bar(aes(fill = condition), stat = "identity", position = "dodge",
           alpha = .6) +
  ylim(c(0,6)) +
    labs(title= "Children's Selective Trust of Helpful over Tricky informant", 
       subtitle = "Separated by Condition and Age group", 
       x = "Age Group", y = "Average Trust Score" ) +
     geom_errorbar(aes(ymin = average_trust - 1.96*se_average_trust, 
                       ymax = average_trust + 1.96*se_average_trust,
                       group = condition),
                  width=.2,                    
                  position = position_dodge(width = 0.9)) +
  theme_classic() +
  scale_fill_viridis_d(option = "cividis")

plot1 + theme(axis.title.x = element_text(size=14),
           axis.text.x= element_text(size=12),
           axis.title.y = element_text(size=14),
           axis.text.y= element_text(size=12),
           plot.title = element_text(size=16),
           plot.subtitle = element_text(size=14))
  


#add annotations for group averages, make the title for the general public

plot2 <- ggplot(summary_p, aes(as.factor(agegroup.x), average_trust )) +
  geom_bar(aes(fill = condition), stat = "identity", position = "dodge",
           alpha = .5) +
  ylim(c(0,6)) +
    labs(title= "Only 4-year-olds trust the more reliable informant!", 
       subtitle = "", 
       x = "Age Group", y = "Average Trust Score" ) +
     geom_errorbar(aes(ymin = average_trust - 1.96*se_average_trust, 
                       ymax = average_trust + 1.96*se_average_trust,
                       group = condition),
                  width=.1, 
                  alpha = .5,
                  color = "gray40",
                  position = position_dodge(width = 0.9)) +
  annotate("text", label = "3.43", x = .79, y = 3.43, fontface = "bold", size = 4) +
  annotate("text", label = "3.32", x = 1.2, y = 3.32, fontface = "bold", size = 4) +
  annotate("text", label = "4.71", x = 1.8, y = 4.71, fontface = "bold", size = 4) +
  annotate("text", label = "3.96", x = 2.2, y = 3.96, fontface = "bold", size = 4) +
  theme_classic() +
  scale_fill_viridis_d(option = "cividis")


plot2 + theme(axis.title.x = element_text(size=14),
           axis.text.x= element_text(size=12),
           axis.title.y = element_text(size=14),
           axis.text.y= element_text(size=12),
           plot.title = element_text(size=16),
           plot.subtitle = element_text(size=14))

# going to try to take off error bars so it's easier to see means. I tried putting them above the error bars but it looks odd. Let me know if it's not a good idea. 


plot3 <- ggplot(summary_p, aes(as.factor(agegroup.x), average_trust )) +
  geom_bar(aes(fill = condition), stat = "identity", position = "dodge",
           alpha = .5) +
  ylim(c(0,6)) +
    labs(title= "Only 4-year-olds trust the more reliable informant!", 
       subtitle = "Children's trust scores in a sticker-finding game", 
       x = "Age Group", y = "Average Trust Score of Reliable Informant" ) +
  annotate("text", label = "3.43", x = .79, y = 3.61, size = 6) +
  annotate("text", label = "3.32", x = 1.2, y = 3.5, size = 6) +
  annotate("text", label = "4.71", x = 1.8, y = 4.9, size = 6) +
  annotate("text", label = "3.96", x = 2.2, y = 4.1, size = 6) +
  theme_classic() +
  scale_fill_viridis_d(option = "cividis")


plot3 + theme(axis.title.x = element_text(size=14),
           axis.text.x= element_text(size=12),
           axis.title.y = element_text(size=14),
           axis.text.y= element_text(size=12),
           plot.title = element_text(size=18),
           plot.subtitle = element_text(size=14))

# we hear so often that bar plots are so bad so here's an alternative that shows all the data. 
v1 <- v1 %>% 
  mutate(condition = recode(condition,
                "first-hand" = "First-hand",
                "second-hand" = "Second-hand"))

plot3.1 <- ggplot(v1, aes(as.factor(agegroup.x), total_trust, color = (as.factor(agegroup.x)))) +
  geom_point(alpha = 0.9, position = position_jitter(height = .2, width = .2)) +
  facet_wrap(~condition) +
  theme_minimal() +
  scale_color_brewer(palette = "Accent") +
    labs(title= "Only 4-year-olds trust the more reliable informant!", 
       subtitle = "Children's trust scores in a sticker-finding game", 
       x = "Age Group",
       y = "Trust Score of Reliable Informant")

plot3.1 +
  theme(axis.title.x = element_text(size=14),
           axis.text.x= element_text(size=12),
           axis.title.y = element_text(size=14),
           axis.text.y= element_text(size=12),
           plot.title = element_text(size=15),
           plot.subtitle = element_text(size=14),
           strip.text.x = element_text(size = 12),
           legend.position = "none")
  

#feedback requests:

# I would love feedback on the colors. I am trying to find some that are colorblind friendly but I think they all look ugly.

# Daniel suggested adding annotaions, does it help to have the means there? It's look awkward to me because of the error bars, but seems odd to take them off. 

# Is text big enough? Would this be understood by a general audience?

#let me know thoughts on plot 3.1

```


Viz 2 - aimed at scientific community

Message: it seems that there is a positive relationship between four-year-olds, who received first-hand information regarding informants' traits, performance on the training trials and trust of more reliable informant. This relationship does not seem evident for 3-year-olds. 
```{r viz 2, fig.width = 5, fig.height = 6}
df3 <- df2 %>% 
  filter(condition == "1") 

#Let's make this sad scatter plot better

ggplot(df3, aes(trainingtotal, total_trust)) +
  geom_point()

#theme minimal added jitter for overlapping points 
ggplot(df3, aes(trainingtotal, total_trust)) +
  geom_point(position = "jitter") +
  geom_smooth(method = "lm", fullrange = TRUE) +
  theme_minimal() +
  labs(title= "Children's Performance on Training Trials Related \nto Total Trust of Reliable Informant", 
       subtitle = "", 
       x = "Training Total", y = "Trust Score" )

#color dots and regression line by age group 
plot <- ggplot(df3, aes(trainingtotal, total_trust, colour = as.factor(agegroup.x))) +
  geom_point(size = 2) +
  geom_jitter(position = position_jitter(height = .6, width = 0)) +
  geom_smooth(method = "lm", fullrange = TRUE, se = FALSE) +
  scale_color_brewer(palette = "Accent",
                     name = "Age Group",
                     breaks = c("3", "4"),
                     labels = c("3-year-olds", "4-year-olds")) +
  theme_minimal() +
  labs(title= "Children's Performance on Training Trials Related to Total \nTrust of Reliable Informant", 
       subtitle = "Relationship separated by age", 
       x = "Training Total",
       y = "Trust Score" ) +
  annotate("text", label = "3-year-olds", x = 2.8, y = 3.1, size = 4.5) +
  annotate("text", label = "4-year-olds", x = 2.8, y = 5.1, size = 4.5) +
  guides(color = "none")

plot + theme(axis.title.x = element_text(size=14),
           axis.text.x= element_text(size=12),
           axis.title.y = element_text(size=14),
           axis.text.y= element_text(size=12),
           plot.title = element_text(size=15),
           plot.subtitle = element_text(size=14))
  
#change colors, consider not using coord equal, find colors that are color blind safe and not ugly like the viridis ones 

```

Viz 3 - aimed at scientific community

message: it seems that parents of 4-year-olds more accurately report on their 4-year-olds executive function (EF) performance. Parent reports of 3-year-olds EF seem to overestimate their performance. 
```{r viz 3}

#make data into long format
df4 <- data %>%
  gather(BRIEF_item, BRIEF_response, "BRIEF1":"BRIEF63") %>%
  mutate(BRIEF_item = as.factor(parse_number(BRIEF_item))) %>%
  mutate(BRIEF_response = as.numeric(BRIEF_response)) %>% 
  group_by(subid) %>%  
  mutate(child_EF = ((flanker_rs + dccs_rs)/2)) %>% 
  mutate(parent_EF = sum(BRIEF_response)) %>% 
  filter(!is.na(parent_EF)) %>% 
  filter(!is.na(child_EF)) %>% 
  mutate(parent_EF_RS = (63 - parent_EF)) %>% 
  mutate(agegroup.x = recode(agegroup.x,
                              "3" = "3-year-olds",
                              "4" = "4-year-olds")) 

#checking for outliers for each variable 

boxplot(df4$child_EF)
boxplot(df4$parent_EF_RS)
ggplot(df4, aes(child_EF, parent_EF_RS, group = agegroup.x)) + 
  geom_boxplot()

#my plan was to highlight outliers in the scatterplot, as you can see there are no outliers. may do something else! Suggestions to make it look fancier?

plot4 <- ggplot(df4, aes(child_EF, parent_EF_RS)) +
  geom_point(size = 1.5, color = "gray40") +
  geom_smooth(method = "lm") +
  facet_wrap(~agegroup.x) +
  theme_bw() + 
  labs(title= "Children's Executive Function Score in Relation to Parent Report \nof Executive Function",
       subtitle = "Relationship separated by age", 
       x = "Children's Executive Function Score",
       y = "Parent-Report of Executive Function" )

plot4 + theme(axis.title.x = element_text(size=14),
           axis.text.x= element_text(size=12),
           axis.title.y = element_text(size=14),
           axis.text.y= element_text(size=12),
           plot.title = element_text(size=15),
           plot.subtitle = element_text(size=14),
           strip.text.x = element_text(size = 12))


```

viz 4: aimed at scientific community 

message: children's ability to trust a reliable informant over an unreliable informant seems to increase with age. 
```{r}
df5 <- df2 %>% 
mutate(gender = recode(gender,
                "0" = "Girls",
                "1" = "Boys"))

plot5 <- ggplot(df5, aes(exactage.x, total_trust)) +
  geom_point(alpha = 0.6) +
  facet_wrap(~gender) +
  gghighlight(total_trust > 1) +
  geom_smooth(method = "lm") +
  theme_bw() +
  labs(title= "Children's Age in Relation to Trust of the Reliable Informant",
       subtitle = "Separated by Gender", 
       x = "Age in Months",
       y = "Total Trust Score") 

plot5 + theme(axis.title.x = element_text(size=14),
           axis.text.x= element_text(size=12),
           axis.title.y = element_text(size=14),
           axis.text.y= element_text(size=12),
           plot.title = element_text(size=15),
           plot.subtitle = element_text(size=14),
           strip.text.x = element_text(size = 12),
           panel.spacing.x = unit(1.5, "cm"))



```


