---
title: "Data Vizualizations Portfolio"
author: "Karlena Ochoa"
date: "February 24, 2019"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r load libraries }

library(tidyverse)
library(janitor)
library(rio)
library(magrittr)
library(here)
library(knitr)
library(stats)
library(gghighlight)
#devtools::install_github("yixuan/showtext")
library(showtext)
library(ggthemes)
library(grid)

```


```{r import data}

df <- import(here("data", "BREIF_Data_100p_final_proj.csv"),
             na.string = "999")
             
df2 <- import(here("data", "data_trust_final_proj.csv"),
              na.string = "999")

# turn 999 into nas

#join data and select variables of interest

data <- left_join(df2, df, by = "subid") %>% 
  select(1, 4:7, 12:23, 25:28, 33:44, 56:73, 81:143)
          
```


Viz 1: Aimed at general community!

Message: only 4-year-olds trust the reliable informant over the unreliable informant, even when 3-year-olds receive first-hand information they are equally choosing the reliable (helpful) and unreliable informant (tricky) in a sticker-finding task. 
```{r viz 1}

#mutate variables to factors and label, make average trust variable
v1 <- data %>% 
mutate(condition = as.factor(ifelse(condition == 1, "first-hand",     
                                      "second-hand"))) %>%
  mutate(gender = as.factor(ifelse(gender == 1, "male", "female"))) %>%
  group_by(subid) %>%
  mutate(average_trust = mean(total_trust))

#AAC: I really like your use of the ifelse command in the mutate while factoring- I never think to do this, but I think it's an awesome tool! Definitely going to use it in the future

summary_p <- v1 %>%
  group_by(agegroup.x, condition) %>%
  summarise(n = n(),
            average_trust = mean(total_trust),
            sd_average_trust = sd(total_trust),
            se_average_trust = sd_average_trust / (sqrt(n)))

plot0 <- ggplot(summary_p, aes(as.factor(agegroup.x), average_trust)) +
  geom_bar(aes(fill = condition), stat = "identity", position = "dodge",
           alpha = .7) +
    labs(title= "Children's Selective Trust of Helpful over Tricky informant", 
       subtitle = "Separated by Condition and Age group", 
       x = "Age Group", y = "Average Trust Score" ) +
     geom_errorbar(aes(ymin = average_trust - 1.96*se_average_trust, 
                       ymax = average_trust + 1.96*se_average_trust,
                       group = condition),
                  width=.2,                    
                  position = position_dodge(width = 0.9)) +
  theme_minimal() 
#AAC: Because is aimed at the general community rather than the scientific community, I feel like error bars might not be the best fit for this audience (and I see you eliminated them in the below visualizations :D)
plot0 

#improve plot by changing colors, make text bigger 

plot1 <- ggplot(summary_p, aes(as.factor(agegroup.x), average_trust)) +
  geom_bar(aes(fill = condition), stat = "identity", position = "dodge",
           alpha = .6) +
  ylim(c(0,6)) +
    labs(title= "Children's Selective Trust of Helpful over Tricky informant", 
       subtitle = "Separated by Condition and Age group", 
       x = "Age Group", y = "Average Trust Score" ) +
     geom_errorbar(aes(ymin = average_trust - 1.96*se_average_trust, 
                       ymax = average_trust + 1.96*se_average_trust,
                       group = condition),
                  width=.2,                    
                  position = position_dodge(width = 0.9)) +
  theme_classic() +
  scale_fill_viridis_d(option = "cividis")


plot1 + theme(axis.title.x = element_text(size=14),
           axis.text.x= element_text(size=12),
           axis.title.y = element_text(size=14),
           axis.text.y= element_text(size=12),
           plot.title = element_text(size=16),
           plot.subtitle = element_text(size=14))
#AAC: Love the improved colors. Because you used theme_classic instead of theme_minimal, and thus eliminated the gridlines in the background of this visualization, I think it might not be necessary to use alpha to make the bars transparent because there's nothing really to see behind them. And the color is probably nicer without the alpha
  


#add annotations for group averages, make the title for the general public

plot2 <- ggplot(summary_p, aes(as.factor(agegroup.x), average_trust )) +
  geom_bar(aes(fill = condition), stat = "identity", position = "dodge",
           alpha = .5) +
  ylim(c(0,6)) +
    labs(title= "Only 4-year-olds trust the more reliable informant!", 
       subtitle = "", 
       x = "Age Group", y = "Average Trust Score" ) +
     geom_errorbar(aes(ymin = average_trust - 1.96*se_average_trust, 
                       ymax = average_trust + 1.96*se_average_trust,
                       group = condition),
                  width=.1, 
                  alpha = .5,
                  color = "gray40",
                  position = position_dodge(width = 0.9)) +
  annotate("text", label = "3.43", x = .79, y = 3.43, fontface = "bold", size = 4) +
  annotate("text", label = "3.32", x = 1.2, y = 3.32, fontface = "bold", size = 4) +
  annotate("text", label = "4.71", x = 1.8, y = 4.71, fontface = "bold", size = 4) +
  annotate("text", label = "3.96", x = 2.2, y = 3.96, fontface = "bold", size = 4) +
  theme_classic() +
  scale_fill_viridis_d(option = "cividis")


plot2 + theme(axis.title.x = element_text(size=14),
           axis.text.x= element_text(size=12),
           axis.title.y = element_text(size=14),
           axis.text.y= element_text(size=12),
           plot.title = element_text(size=16),
           plot.subtitle = element_text(size=14))
#AAC: Yeah, I can see how the error bars and annotations might be too much together. Go with your gut :D 

# going to try to take off error bars so it's easier to see means. I tried putting them above the error bars but it looks odd. Let me know if it's not a good idea. 


plot3 <- ggplot(summary_p, aes(as.factor(agegroup.x), average_trust )) +
  geom_bar(aes(fill = condition), stat = "identity", position = "dodge",
           alpha = .5) +
  ylim(c(0,6)) +
    labs(title= "Only 4-year-olds trust the more reliable informant!", 
       subtitle = "Children's trust scores in a sticker-finding game", 
       x = "Age Group", y = "Average Trust Score of Reliable Informant" ) +
  annotate("text", label = "3.43", x = .79, y = 3.61, size = 6) +
  annotate("text", label = "3.32", x = 1.2, y = 3.5, size = 6) +
  annotate("text", label = "4.71", x = 1.8, y = 4.9, size = 6) +
  annotate("text", label = "3.96", x = 2.2, y = 4.1, size = 6) +
  theme_classic() +
  scale_fill_viridis_d(option = "cividis")


plot3 + theme(axis.title.x = element_text(size=14),
           axis.text.x= element_text(size=12),
           axis.title.y = element_text(size=14),
           axis.text.y= element_text(size=12),
           plot.title = element_text(size=18),
           plot.subtitle = element_text(size=14))

#AAC: I like the title and I think it's awesome for the general public! One thing you might consider is shortening the y axis title, capitalizing the "Condition" in the legend, and making it clear that the age groups are 3 and 4 years*. 

# we hear so often that bar plots are so bad so here's an alternative that shows all the data. 
v1 <- v1 %>% 
  mutate(condition = recode(condition,
                "first-hand" = "First-hand",
                "second-hand" = "Second-hand"))

plot3.1 <- ggplot(v1, aes(as.factor(agegroup.x), total_trust, color = (as.factor(agegroup.x)))) +
  geom_point(alpha = 0.9, position = position_jitter(height = .2, width = .2)) +
  facet_wrap(~condition) +
  theme_minimal() +
  scale_color_brewer(palette = "Accent") +
    labs(title= "Only 4-year-olds trust the more reliable informant!", 
       subtitle = "Children's trust scores in a sticker-finding game", 
       x = "Age Group",
       y = "Trust Score of Reliable Informant")

plot3.1 +
  theme(axis.title.x = element_text(size=14),
           axis.text.x= element_text(size=12),
           axis.title.y = element_text(size=14),
           axis.text.y= element_text(size=12),
           plot.title = element_text(size=15),
           plot.subtitle = element_text(size=14),
           strip.text.x = element_text(size = 12),
           legend.position = "none") #I really like how you separated out your plot code from your theme edits like this. It made it really easy to understand!

#AAC: I love this one! I always like looking at the raw data points more than summary statistics, and I really like jitter plots! Maybe you could add an annotation or a line on this plot marking the means for each group, which might help for comparisons? 
#Also my go-to color is always the original viridis but that's just personal preference
#I think the text is definitely big enough and I think that (minus perhaps the error bars) that this can be understood by a general audience. Awesome job! 
  

#feedback requests:

# I would love feedback on the colors. I am trying to find some that are colorblind friendly but I think they all look ugly.

# Daniel suggested adding annotaions, does it help to have the means there? It's look awkward to me because of the error bars, but seems odd to take them off. 

# Is text big enough? Would this be understood by a general audience?

#let me know thoughts on plot 3.1

```


Viz 2 - aimed at scientific community

Message: it seems that there is a positive relationship between four-year-olds, who received first-hand information regarding informants' traits, performance on the training trials and trust of more reliable informant. This relationship does not seem evident for 3-year-olds. 
```{r viz 2, fig.width = 5, fig.height = 6}
df3 <- df2 %>% 
  filter(condition == "1") 

#Let's make this sad scatter plot better

ggplot(df3, aes(trainingtotal, total_trust)) +
  geom_point()

#theme minimal added jitter for overlapping points 
ggplot(df3, aes(trainingtotal, total_trust)) +
  geom_point(position = "jitter") +
  geom_smooth(method = "lm", fullrange = TRUE) +
  theme_minimal() +
  labs(title= "Children's Performance on Training Trials Related \nto Total Trust of Reliable Informant", 
       subtitle = "", 
       x = "Training Total", y = "Trust Score" )
#AAC: Awesome use of jitter! 

#color dots and regression line by age group 
plot <- ggplot(df3, aes(trainingtotal, total_trust, colour = as.factor(agegroup.x))) +
  geom_point(size = 2) +
  geom_jitter(position = position_jitter(height = .6, width = 0)) +
  geom_smooth(method = "lm", fullrange = TRUE, se = FALSE) +
  scale_color_brewer(palette = "Accent",
                     name = "Age Group",
                     breaks = c("3", "4"),
                     labels = c("3-year-olds", "4-year-olds")) +
  theme_minimal() +
  labs(title= "Children's Performance on Training Trials Related to Total \nTrust of Reliable Informant", 
       subtitle = "Relationship separated by age", 
       x = "Training Total",
       y = "Trust Score" ) +
  annotate("text", label = "3-year-olds", x = 2.8, y = 3.1, size = 4.5) +
  annotate("text", label = "4-year-olds", x = 2.8, y = 5.1, size = 4.5) +
  guides(color = "none")

plot + theme(axis.title.x = element_text(size=14),
           axis.text.x= element_text(size=12),
           axis.title.y = element_text(size=14),
           axis.text.y= element_text(size=12),
           plot.title = element_text(size=15),
           plot.subtitle = element_text(size=14))
#Great use of annotations!
  
#change colors, consider not using coord equal, find colors that are color blind safe and not ugly like the viridis ones 



#AAC:Awesome! Great use of linear modeling, and I always like to see the underlying data points like you have. I also thought it was smart to not depiect the se in the linear model, because you are already displaying the data points and it might be too much going on. Are some of these points different sizes? That's a little bit confusing to me. I would also suggest changing the colors to one of the color blind palettes, just because these seem a little faint. Also, I'm curious why you set the size for this plot but not the previous plot. All super small comments, but overall I think the plot and audience are very appropriate! 

```

Viz 3 - aimed at scientific community

message: it seems that parents of 4-year-olds more accurately report on their 4-year-olds executive function (EF) performance. Parent reports of 3-year-olds EF seem to overestimate their performance. 
```{r viz 3}

#make data into long format
df4 <- data %>%
  gather(BRIEF_item, BRIEF_response, "BRIEF1":"BRIEF63") %>%
  mutate(BRIEF_item = as.factor(parse_number(BRIEF_item))) %>%
  mutate(BRIEF_response = as.numeric(BRIEF_response)) %>% 
  group_by(subid) %>%  
  mutate(child_EF = ((flanker_rs + dccs_rs)/2)) %>% 
  mutate(parent_EF = sum(BRIEF_response)) %>% 
  filter(!is.na(parent_EF)) %>% 
  filter(!is.na(child_EF)) %>% 
  mutate(parent_EF_RS = (63 - parent_EF)) %>% 
  mutate(agegroup.x = recode(agegroup.x,
                              "3" = "3-year-olds",
                              "4" = "4-year-olds")) 
#AAC:Great use of spread and tidying

#checking for outliers for each variable 

boxplot(df4$child_EF)
boxplot(df4$parent_EF_RS)
ggplot(df4, aes(child_EF, parent_EF_RS, group = agegroup.x)) + 
  geom_boxplot()

#my plan was to highlight outliers in the scatterplot, as you can see there are no outliers. may do something else! Suggestions to make it look fancier?

#AAC: That would've been cool, bummer it didn't work out ): but also good that the data has no outliers :D 

plot4 <- ggplot(df4, aes(child_EF, parent_EF_RS)) +
  geom_point(size = 1.5, color = "gray40") +
  geom_smooth(method = "lm") +
  facet_wrap(~agegroup.x) +
  theme_bw() + 
  labs(title= "Children's Executive Function Score in Relation to Parent Report \nof Executive Function",
       subtitle = "Relationship separated by age", 
       x = "Children's Executive Function Score",
       y = "Parent-Report of Executive Function" )

plot4 + theme(axis.title.x = element_text(size=14),
           axis.text.x= element_text(size=12),
           axis.title.y = element_text(size=14),
           axis.text.y= element_text(size=12),
           plot.title = element_text(size=15),
           plot.subtitle = element_text(size=14),
           strip.text.x = element_text(size = 12))

#AAC: I really like the look of this faceted plot. Simple, but effective. My only question is that you said that "parents of 4-year-olds more accurately report on their 4-year-olds executive function (EF) performance. Parent reports of 3-year-olds EF seem to overestimate their performance", but despite this pattern, the linear models look kind of similar, despite having somewhat different intercepts, and seem to have similar slopes. I can see the relationship that you are talking about by looking at the scatterplot of points. Perhaps you could display this relationship as the average dfference between parent-reported scores and actual reported scores for the two age groups? Or highlight the specific details of the plot that display the relationship that you're talking about (also it could be that I just don't understand enough about stats :D )
```

viz 4: aimed at scientific community 

message: children's ability to trust a reliable informant over an unreliable informant seems to increase with age. 
```{r}
df5 <- df2 %>% 
mutate(gender = recode(gender,
                "0" = "Girls",
                "1" = "Boys"))

plot5 <- ggplot(df5, aes(exactage.x, total_trust)) +
  geom_point(alpha = 0.6) +
  facet_wrap(~gender) +
  gghighlight(total_trust > 1) +
  geom_smooth(method = "lm") +
  theme_bw() + #I like your experimentation with different themes
  labs(title= "Children's Age in Relation to Trust of the Reliable Informant",
       subtitle = "Separated by Gender", 
       x = "Age in Months",
       y = "Total Trust Score") 

plot5 + theme(axis.title.x = element_text(size=14),
           axis.text.x= element_text(size=12),
           axis.title.y = element_text(size=14),
           axis.text.y= element_text(size=12),
           plot.title = element_text(size=15),
           plot.subtitle = element_text(size=14),
           strip.text.x = element_text(size = 12),
           panel.spacing.x = unit(1.5, "cm")) # I love this panel spacing. I didn't know you could do that, and I think it looks great!

#AAC: Again, I really like the look of these faceted linear model plots! I'm a little confused why you highlighted the scores specifically above 1, but it may just be my unfamiliarity with your data. Making that more clear woluld help me better interpret the plot.


# Great job overall on all your plots, Karlena! You even included a bonus one- I think you're in an awesome spot!

```


